---
site: workflowr::wflow_site
params:
  dt.exprs: dt.exprs
  dt.cytnum: dt.cytnum
  cytokine_nodes: cytokine_nodes
  markers: markers
  set_title: report_title
  set_author: report_author
title: "`r params$set_title`"
author: "`r params$set_author`"
date: "`r Sys.Date()`"
always_allow_html: true
editor_options: 
  chunk_output_type: console
output: html_document 
runtime: shiny
---

<style>
body{text-align: justify}
pre code, pre, code {
  white-space: pre !important;
  overflow-x: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
</style>

Authors: `r params$set_author`

```{r global_options, echo=FALSE, message=FALSE, warning=FALSE}
###--- Library
library(data.table)
library(rmarkdown)
library(knitr)
library(here)
library(tidyverse)
```

The purpose of this QC report is to assess the quality of the data before running clustering algorithms. Summary tables accounting for samples and unreliable data are blinded to treatment group.  
    

# Summary tables of samples and readouts
________________________________________

## Data Input

### Table 1 | Listing of columns (exprs)

List of columns including pData, FlowJo gates and FI.  

```{r table_1, echo=FALSE, message=FALSE, warning=FALSE, results="asis"}
#- Summary
rmarkdown::paged_table(dt.exprs, options = list(rows.print = 10))
```

### Table 2 | Listing of columns (CYTNUM & NSUB)

List of columns including NSUB and CYTNUM.  

```{r table_2, echo=FALSE, message=FALSE, warning=FALSE, results="asis"}
#- Summary
rmarkdown::paged_table(dt.cytnum %>% filter(boolean_CYTNUM == TRUE) %>% select(-c(boolean_CYTNUM)), options = list(rows.print = 10))
```

## Summary

### Table 3 | Sample summary by BATCH

Number of samples analyzed, by BATCH.  

```{r table_3, echo=FALSE, message=FALSE, warning=FALSE, results="asis"}
#- Summary
if("reliable_flag" %in% colnames(dt.exprs)) {
  dt.exprs %>%
    dplyr::group_by(BATCH, reliable_flag) %>%
    dplyr::summarise(n = length(unique(FCS))) -> tab
}else{
  dt.exprs %>%
    dplyr::group_by(BATCH) %>%
    dplyr::summarise(n = length(unique(FCS))) -> tab
}
rmarkdown::paged_table(tab, options = list(rows.print = 5))
```

### Table 4 | Number of cells, by BATCH

Number of cells analyzed, by BATCH.  

```{r table_4, echo=FALSE, message=FALSE, warning=FALSE, results="asis"}
#- Summary
if("reliable_flag" %in% colnames(dt.exprs)) {
  dt.exprs %>%
    dplyr::filter(reliable_flag == 1) %>%
    dplyr::group_by(BATCH) %>%
    dplyr::summarise(n = n()) -> tab
}else{
  dt.exprs %>%
    dplyr::group_by(BATCH) %>%
    dplyr::summarise(n = n()) -> tab
}
rmarkdown::paged_table(tab, options = list(rows.print = 5))
```

### Table 5 | Number of cells, by SAMPLE

Number of cells analyzed, by SAMPLE (.FCS)  

```{r table_5, echo=FALSE, message=FALSE, warning=FALSE, results="asis"}
#- Summary
if("reliable_flag" %in% colnames(dt.exprs)) {
  dt.exprs %>%
    dplyr::filter(reliable_flag == 1) %>%
    dplyr::group_by(FCS) %>%
    dplyr::summarise(n = n()) -> tab
}else{
  dt.exprs %>%
    dplyr::group_by(FCS) %>%
    dplyr::summarise(n = n()) -> tab
}
rmarkdown::paged_table(tab, options = list(rows.print = 5))
```
    

# Nb. of cytokines and/or functional activation markers
_______________________________________________________

## Total of cells expressing cytokines and/or functional activation markers

```{r plot_1, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=5}
###--- Barplot
if("reliable_flag" %in% colnames(dt.exprs)) {
  dt.ggplot <- dt.exprs %>%
    dplyr::filter(reliable_flag == 1) %>%
    dplyr::group_by(CYTNUM) %>%
    dplyr::summarize(n = n())
}else{
  dt.ggplot <- dt.exprs %>%
    dplyr::group_by(CYTNUM) %>%
    dplyr::summarize(n = n())
}
dt.ggplot %>%
  ggplot(aes(x = as.character(CYTNUM), y = n)) +
    geom_bar(aes(fill = as.character(CYTNUM)), stat = "identity") +
    geom_label(aes(label = n), vjust = 1, color = "black", size = 3) +
    scale_y_continuous(trans = "sqrt") +
    scale_fill_manual(values = colorRamps::matlab.like2(n = length(unique(dt.ggplot$CYTNUM)))) +
    labs(x = "Nb. cytokines and/or functional activation markers", y = "n") +
    theme_bw() +
    theme(legend.position = "none", 
          panel.grid = element_blank(),
          axis.title = element_text(size = 13),
          axis.text = element_text(size = 11))
```

## Pie charts - For each cytokine, % of co-expression (1, 2, >=3)

```{r plot_2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=4.5, fig.height=4.5}
###--- Barplot
if(!is.null(cytokine_nodes)){
  for(i in cytokine_nodes) {
    if("reliable_flag" %in% colnames(dt.exprs)) {
      dt.ggplot <- dt.exprs %>%
        dplyr::filter(reliable_flag == 1) %>%
        dplyr::filter(get({{i}}) == TRUE) %>%
        dplyr::select(CYTNUM, i) %>%
        dplyr::mutate(CYTNUM = case_when(CYTNUM >= 3 ~ ">=3", .default = as.character(CYTNUM))) %>%
        dplyr::group_by(CYTNUM) %>%
        dplyr::summarize(n = n())
    }else{
      dt.ggplot <- dt.exprs %>%
        dplyr::filter(get({{i}}) == TRUE) %>%
        dplyr::select(CYTNUM, i) %>%
        dplyr::mutate(CYTNUM = case_when(CYTNUM >= 3 ~ ">=3", .default = as.character(CYTNUM))) %>%
        dplyr::group_by(CYTNUM) %>%
        dplyr::summarize(n = n())
    }
    sum <- sum(dt.ggplot$n)
    dt.ggplot <- dt.ggplot %>%
      dplyr::mutate(percent = round(n / sum * 100, 2)) %>%
      dplyr::mutate(CYTNUM = factor(x = CYTNUM, levels = c("1", "2", ">=3"))) %>%
      dplyr::arrange(desc(CYTNUM))
    dt.ggplot %>%
      ggplot(aes(x = "", y = percent)) +
        geom_bar(aes(fill = CYTNUM), stat = "identity", width = 1) +
        geom_label(aes(label = percent),
                   position = position_stack(vjust = 0.5),
                   show.legend = FALSE) +
        coord_polar("y", start = 0) +
        scale_fill_viridis_d() +
        labs(x = NULL, y = NULL, title = i) + 
        theme_bw() +
        theme(plot.title = element_text(size = 8),
              axis.title = element_blank(),
              axis.text = element_blank(),
              axis.ticks = element_blank(),
              panel.grid = element_blank()) -> plot 
    plot %>%
      print()
  }
}
```


# Distribution of transformed FI
________________________________

## biexp

```{r data_structure_1, echo=FALSE, message=FALSE, warning=FALSE, fig.width=4, fig.height=7, results='asis'}
###--- QC
#- if condition
if(!is.null(markers)){
  markers.tmp <- markers[str_detect(string = markers, pattern = "biexp")]
  if(length(markers.tmp) == 0){
    print("No biexp values")
  }
}

#- Ridge plots by BATCH
if(!is.null(markers)){
  markers.tmp <- markers[str_detect(string = markers, pattern = "biexp")]
  for(marker in markers.tmp){
    # data.table()
    if("reliable_flag" %in% colnames(dt.exprs)) {
      dt.ggplot <- dt.exprs %>%
        dplyr::filter(reliable_flag == 1) %>%
        dplyr::select(marker, BATCH)
    }else{
      dt.ggplot <- dt.exprs %>%
        dplyr::select(marker, BATCH)
    }
    colnames(dt.ggplot)[1] <- "value"
    dt.ggplot$marker <- marker
  
    # Figure
    dt.ggplot %>%
      ggplot() +
      ggridges::geom_density_ridges(aes(x = value, y = BATCH, fill = BATCH), bandwidth = .1, lwd = .25) +
      geom_point(aes(x = value, y = BATCH), data = dt.ggplot %>% filter(value < -4500), pch = "|") +
      geom_vline(xintercept = 0, alpha = .5, lwd = .25, lty = 2) +
      scale_x_continuous(limits = c(-75000, 10000)) +
      scale_fill_viridis_d(option = "A", begin = .2) +
      facet_wrap(~ marker) +
      theme_bw() +
      labs(x = NULL, y = NULL) +
      theme(legend.position = "none", 
            axis.title = element_text(size = 16), 
            panel.grid = element_blank(), 
            strip.text = element_text(size = 16)) -> plot
    print(plot)
  }
}
```

## asinh

```{r data_structure_2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=4, fig.height=7, results='asis'}
###--- QC
#- if condition
if(!is.null(markers)){
  markers.tmp <- markers[str_detect(string = markers, pattern = "asinh")]
  markers.tmp <- markers.tmp[!str_detect(string = markers.tmp, pattern = "asym")]
  if(length(markers.tmp) == 0){
    print("No asinh values")
  }
}  

#- Ridge plots by BATCH
if(!is.null(markers)){
  markers.tmp <- markers[str_detect(string = markers, pattern = "asinh")]
  markers.tmp <- markers.tmp[!str_detect(string = markers.tmp, pattern = "asym")]
  for(marker in markers.tmp){
    # data.table()
    if("reliable_flag" %in% colnames(dt.exprs)) {
      dt.ggplot <- dt.exprs %>%
        dplyr::filter(reliable_flag == 1) %>%
        dplyr::select(marker, BATCH)
    }else{
      dt.ggplot <- dt.exprs %>%
        dplyr::select(marker, BATCH)
    }
    colnames(dt.ggplot)[1] <- "value"
    dt.ggplot$marker <- marker
  
    # Figure
    dt.ggplot %>%
      ggplot() +
      ggridges::geom_density_ridges(aes(x = value, y = BATCH, fill = BATCH), bandwidth = .1, lwd = .25) +
      geom_point(aes(x = value, y = BATCH), data = dt.ggplot %>% filter(value < -3), pch = "|") +
      geom_vline(xintercept = 0, alpha = .5, lwd = .25, lty = 2) +
      scale_x_continuous(limits = c(-7.5, 7.5)) +
      scale_fill_viridis_d(option = "A", begin = .2) +
      facet_wrap(~ marker) +
      theme_bw() +
      labs(x = NULL, y = NULL) +
      theme(legend.position = "none", 
            axis.title = element_text(size = 16), 
            panel.grid = element_blank(), 
            strip.text = element_text(size = 16)) -> plot
    print(plot)
  }
}
```

## asinh + asym

```{r data_structure_3, echo=FALSE, message=FALSE, warning=FALSE, fig.width=4, fig.height=7, results='asis'}
###--- QC
#- if condition
if(!is.null(markers)){
  markers.tmp <- markers[str_detect(string = markers, pattern = "asinh_asym")]
  if(length(markers.tmp) == 0){
    print("No asinh+asym values")
  }
}

#- Ridge plots by BATCH
if(!is.null(markers)){
  markers.tmp <- markers[str_detect(string = markers, pattern = "asinh_asym")]
  for(marker in markers.tmp){
    # data.table()
    if("reliable_flag" %in% colnames(dt.exprs)) {
      dt.ggplot <- dt.exprs %>%
        dplyr::filter(reliable_flag == 1) %>%
        dplyr::select(marker, BATCH)
    }else{
      dt.ggplot <- dt.exprs %>%
        dplyr::select(marker, BATCH)
    }
    colnames(dt.ggplot)[1] <- "value"
    dt.ggplot$marker <- marker
  
    # Figure
    dt.ggplot %>%
      ggplot() +
      ggridges::geom_density_ridges(aes(x = value, y = BATCH, fill = BATCH), bandwidth = .1, lwd = .25) +
      geom_point(aes(x = value, y = BATCH), data = dt.ggplot %>% filter(value < -0.236068), pch = "|") +
      geom_vline(xintercept = 0.5857864, alpha = .5, lwd = .25, lty = 2) +
      scale_x_continuous(limits = c(-1, 7.5)) +
      scale_fill_viridis_d(option = "A", begin = .2) +
      facet_wrap(~ marker) +
      theme_bw() +
      labs(x = NULL, y = NULL) +
      theme(legend.position = "none", 
            axis.title = element_text(size = 16), 
            panel.grid = element_blank(), 
            strip.text = element_text(size = 16)) -> plot
    print(plot)
  }
}  
```

